### 2025-05-15 article/article.qmd
first commit

---
title: "The Cost of Exclusion: How Political Discrimination Undermines Trust in Political Institutions in the European Union"
format:
  pdf:
    toc: true
    number-sections: true
    colorlinks: true
editor: source
fig-cap-location: top
fontfamily: inter
fontfamilyoptions: 
  - sfdefault
---

![Discrimination Incidence By Association Ground]("../tables/tab_dicg.png"){#tbl-dicg}

![Demographic Composition Of People Who Has Experienced Political Discrimination or Harrasment]("../tables/tab_dpod.png"){#tbl-dpod}

![Conditional Likelihood Of Demographic Traits and Experiences of Political Discrimination or Harrasment]("../tables/tab_dpod_logit.png"){#tbl-dpodlogit}