### 2025-06-18 article/tabs/descriptive_stats.tex

\begin{table}[hbtp]
\centering
\caption{Sample Demographics and Political Preferences}
\label{table:1}
\footnotesize
\vspace{2mm}
\begin{talltblr}[         %% tabularray outer open
label=none,
note{†} = {Classified as "constrained" if respondents mentioned that their income is not enough fo basic necessities or clothing and "unconstrained" otherwise.},
note{‡} = {Individuals had to answer if they were very interested, interested, a little interested, or not at all interested in politics.},
note{§} = {Political alignment was captured using a vote intention were the respondents had to pick a political parties to vote for if the general elections were happening that weekend.},
remark{Source} = {Eurovoices General Population Poll 2024},
]                     %% tabularray outer close
{                     %% tabularray inner open
colspec={Q[]Q[]Q[]Q[]Q[]Q[]},
hline{4}={1,2,3,4,5,6}{solid, black, 0.1em},
}                     %% tabularray inner close
\toprule
& Unique & Mean & SD & Min & Max \\ \midrule %% TinyTableHeader
Age & 79 & 49.7 & 16.6 & 18.0 & 99.0 \\
Political Ideology & 11 & 5.4 & 2.3 & 0.0 & 10.0 \\
&  & N & \% &  &  \\
Sex & Female & 22,659 & 51.2 &  &  \\
& Male & 21,576 & 48.8 &  &  \\
Area of Residence & Urban & 31,993 & 72.3 &  &  \\
& Rural & 12,242 & 27.7 &  &  \\
Employment & Employed & 24,821 & 56.1 &  &  \\
& Unemployed/Inactive & 19,414 & 43.9 &  &  \\
Citizenship Status & Citizen & 43,224 & 97.7 &  &  \\
& Foreigner & 1,011 & 2.3 &  &  \\
Marital Status & Married & 26,838 & 60.7 &  &  \\
& Not married & 17,397 & 39.3 &  &  \\
Ethnic Group & Ethnic Majority & 26,629 & 60.2 &  &  \\
& Ethnic Minority & 17,606 & 39.8 &  &  \\
Financial Situation\TblrNote{†} & Constrained & 10,103 & 22.8 &  &  \\
& Unconstrained & 34,132 & 77.2 &  &  \\
Education & Higher Education & 14,278 & 32.3 &  &  \\
& No Higher Education & 29,957 & 67.7 &  &  \\
Interest in Politics\TblrNote{‡} & Interested & 21,887 & 49.5 &  &  \\
& Uninterested & 22,348 & 50.5 &  &  \\
Political Alignment\TblrNote{§} & Incumbent Political Party & 7,742 & 17.5 &  &  \\
& Non-Incumbent Political Party & 36,493 & 82.5 &  &  \\
Political D/H & D/H Experience & 7,002 & 15.8 &  &  \\
& No D/H Experience & 37,233 & 84.2 &  &  \\
\bottomrule
\end{talltblr}
\end{table}