### 2025-06-26 article/tabs/het_effects_dissonance.tex

\begin{table}[hbtp]
\centering
\caption{Effect of Political D/H Experiences on Institutional Trust by Level of Political Dissonance and Political Alignment}
\label{table:8}
\footnotesize
\vspace{2mm}
\begin{talltblr}[
label=none,
remark{Note} = {Table shows the estimated coefficient of having experienced political discrimination or harrasment (D/H) on the level of trust in political institutions for four different groups depending on their political alignment and their level of political dissonance. The estimation used the matched sample of 14,004 respondents, region fixed effects, demographic and political covariates, and clustered standard errors at the regional level. *, **, and *** represent statistical significance at $p < 0.05$, $p < 0.01$, and $p < 0.001$, respectively. Estimates shown in the table have been transformed using an exponential transformation to interpret them as a percentage change.},
remark{Source} = {Eurovoices General Population Poll 2024},
]
{                     
colspec={Q[l]Q[l]Q[c]Q[c]Q[c]Q[c]Q[c]},
colsep=4pt
}
\toprule
Pol. Diss. & Pol. Alignment & Estimate & $Pr(>|z|)$ & 95\% CI Lower & 95\% CI Upper & Visualization \\
\midrule
Low & Non-Incumbent & \num{ -0.180}*** & 0.000 & -0.209 & -0.149 & \includegraphics[width=0.47in, height=0.12in]{"viz/heteffects_dissonance_1.png"}\\
 & Incumbent & \num{-0.085} & 0.000 & -0.168 & 0.006 & \includegraphics[width=0.47in, height=0.12in]{"viz/heteffects_dissonance_2.png"}\\
High & Non-Incumbent & \num{-0.211}*** & 0.000 & -0.238 & -0.183 & \includegraphics[width=0.47in, height=0.12in]{"viz/heteffects_dissonance_3.png"}\\
 & Incumbent & \num{-0.179}*** & 0.000 & -0.241 & -0.111 & \includegraphics[width=0.47in, height=0.12in]{"viz/heteffects_dissonance_4.png"}\\
\bottomrule
\end{talltblr}
\end{table}