### 2025-06-26 article/tabs/heterogenous_effects.tex

\begin{table}[hbtp]
\centering
\caption{Effect of Political D/H Experiences on Institutional Trust by Political Alignment}
\label{table:7}
\footnotesize
\vspace{2mm}
\begin{talltblr}[
label=none,
remark{Note} = {Table shows the estimated coefficient of having experienced political discrimination or harrasment (D/H) on the level of trust in political institutions for two different groups depending on their political alignment: non-incumbent and incumbent. The estimation used the matched sample of 14,004 respondents, region fixed effects, demographic and political covariates, and clustered standard errors at the regional level. *, **, and *** represent statistical significance at $p < 0.05$, $p < 0.01$, and $p < 0.001$, respectively. Estimates shown in the table have been transformed using an exponential transformation to interpret them as a percentage change.},
remark{Source} = {Eurovoices General Population Poll 2024},
]
{                     
colspec={Q[l]Q[c]Q[c]Q[c]Q[c]Q[c]},
colsep=4pt
}
\toprule
Pol. Alignment & Estimate & $Pr(>|z|)$ & 95\% CI Lower & 95\% CI Lower & Visualization\\
\midrule
Non-Incumbent &\num{ -0.195}*** & 0.000 & -0.215 & -0.175 & \includegraphics[width=0.47in, height=0.12in]{"viz/heteffects_1.png"}\\
Incumbent & \num{-0.138}*** & 0.000 & -0.189 & -0.084 & \includegraphics[width=0.47in, height=0.12in]{"viz/heteffects_2.png"}\\
\bottomrule
\end{talltblr}
\end{table}